---
title: "Simulations on Recovery Rates"
author: "Yujin Jeong"
date: "6/3/2025"
output: github_document
---

```{r setup, include=FALSE}
setwd("/Users/yujin/Documents/identifying-sparse-treatment-effects")
library(glmnet)
library(stabs)
library(readr)
library(dplyr)
library(tidyverse)
```

This function generates data $\{(X_i, T_i, Y_i)\}_{i=1}^n$ given the sample size $n$, the pre-treatment covariate dimension $m$, the outcome dimension $d$, and the treatment assignment probability $p$. 

```{r}
gen_data <- function(n, m, d, p, alphas, sigma = 1, random_seed=0){
  set.seed(random_seed)
  s = length(alphas)
  # randomly generate model parameters beta and delta
  beta = t(replicate(d, runif(m, min=-1, max=1)))
  delta = t(replicate(s, runif(m)))
  treatment = rbinom(n, 1, p)
  # randomly generate X and Y
  X = matrix(rnorm(n*m, sd=1), nrow=n, ncol=m)
  Y = matrix(0, nrow=n, ncol=d)
  for (i in 1:d){
    if (i <= s) {
      Y[,i] =  X %*% beta[i, ] + treatment * (alphas[i] + X %*% delta[i, ]) + 
        rnorm(n, sd=sigma)
    } else {
      Y[,i] =  X %*% beta[i, ] + rnorm(n, sd=sigma)
    }
  }
  return(list("X" = X, "Y" = Y, "treatment" = treatment))
}
```

The following function performs subset selection using a baseline ranking-based approach combined with stability selection. You can specify the method for pre-treatment covariate adjustment as either "CUPED" or "LIN". The parameters 's_max' and 'cutoff' control the stability selection procedure: 's_max' is the maximum number of variables selected in each subsample, and 'cutoff' is the selection probability threshold for a variable to be retained in the final set. The parameter 'B' specifies the number of subsamples used in the stability selection process.

```{r}
base_method <- function(X, Y, treatment, method='CUPED', s_max=10, cutoff=0.5, B=100){
  
  base_method_helper <- function(X, Y, treatment, alpha=NULL, method='CUPED'){
    
    d = ncol(Y)
    
    if (method == 'CUPED'){
      ## Using CUPED estimator
      lm.fit = lm(Y~X) 
      residuals = lm.fit$residuals
      Y_control = residuals[treatment == 0, ]
      Y_treated = residuals[treatment == 1, ]
      cuped = sapply(1:d, function(i){t.test(Y_treated[,i], Y_control[,i], var.equal=FALSE)$statistic})
      rank_cuped = order(abs(cuped), decreasing = TRUE)
      return (rank_cuped)
    }
    
    if (method == 'LIN') {
      ## Using Lin's estimator
      lm.fit.0 = lm(Y~X, subset = (treatment == 0))
      lm.fit.1 = lm(Y~X, subset = (treatment == 1))
      fitted.0 = predict(lm.fit.0, newdata = as.data.frame(X))
      fitted.1 = predict(lm.fit.1, newdata = as.data.frame(X))
      residuals = Y - mean(treatment) * fitted.0 - mean(1-treatment) * fitted.1
      Y_control = residuals[treatment == 0, ]
      Y_treated = residuals[treatment == 1, ]
      lin = sapply(1:d, function(i){t.test(Y_treated[,i], Y_control[,i], var.equal=FALSE)$statistic})
      rank_lin = order(abs(lin), decreasing = TRUE)
      return (rank_lin)
    }
  }
  
  S = matrix(0, nrow=B, ncol=s_max)
  for (i_b in 1:B){
    indices = sample(1:length(treatment), length(treatment), replace=TRUE)
    X_b = X[indices, ]
    Y_b = Y[indices, ]
    t_b = treatment[indices]
    S[i_b,] = base_method_helper(X_b, Y_b, t_b, method=method)[1:s_max]
  }
  PI = sapply(1:ncol(Y), function(i){sum(S==i)/B})
  return(which(PI>cutoff))
}
```

The following function performs subset selection using the Lasso combined with stability selection. The parameters 's_max' and 'cutoff' control the stability selection procedure: 's_max' is the maximum number of variables selected in each subsample, and 'cutoff' is the selection probability threshold for a variable to be retained in the final set. 

```{r}
our_method <- function(X, Y, treatment, s_max=10, cutoff=0.5){
  penalty = rep(1, ncol(X) + ncol(Y))
  penalty[(ncol(Y)+1):(ncol(X)+ncol(Y))] = 0 # no penalty on X
  weights = (length(treatment)/sum(treatment))^2 * treatment
  weights = weights + (length(treatment)/sum(1-treatment))^2 * (1-treatment)
  selected_set = stabsel(x = scale(cbind(Y, X), center = TRUE, scale = FALSE) * sqrt(weights), 
                         y = (treatment-mean(treatment)) * sqrt(weights),
                         fitfun = glmnet.lasso,
                         args.fitfun = list(family = "gaussian",
                                            alpha = 1, 
                                            nlambda = 200,
                                            penalty.factor = penalty, 
                                            intercept = FALSE, 
                                            standardize = FALSE),
                         cutoff=max(cutoff, 0.8), q=ncol(X)+s_max)$selected
  selected_set = selected_set[!(selected_set %in% ((ncol(Y)+1):(ncol(X)+ncol(Y))))]
  return(selected_set)
}
```

The following code runs simulations to compare the recovery rates of the baseline ranking-based subset selection method and our Lasso-based subset selection method.

```{r, eval=FALSE}
m = 50; d = 500; B = 100

df = data.frame()
n_vec = c(200, 300, 400, 500)
pi_vec = c(0.3, 0.5, 0.7)
alphas = c(1, 1, 1, -1, -1)
exp_vec = seq(from = -1/2, to = -1/5, length.out = 16)
for (pi in pi_vec){
  for (n in n_vec) {
    for (e in exp_vec){
      alpha = 5*(n^(e))
      res = matrix(0, nrow=B, ncol=3)
      sizes = matrix(0, nrow=B, ncol=3)
      for (i_b in 1:B){
        data = gen_data(n, m, d, pi, alphas=alpha*alphas, sigma=1, random_seed=i_b)
        Y = data$Y; X = data$X; treatment = data$treatment
        s1 = base_method(X, Y, treatment, method="CUPED")
        s2 = base_method(X, Y, treatment, method="LIN")
        s3 = our_method(X, Y, treatment)
        res[i_b, ] = c(sum(s1 %in% c(1,2,3,4,5))/5, 
                       sum(s2 %in% c(1,2,3,4,5))/5, 
                       sum(s3 %in% c(1,2,3,4,5))/5)
        sizes[i_b, ] = c(length(s1), length(s2), length(s3))
      }
      sub_df = data.frame(
        "recovery" = colMeans(res),
        "recovery_sd" =  apply(res, 2, sd),
        "recovery_q05" = apply(res, 2, quantile, probs=0.05),
        "recovery_q95" = apply(res, 2, quantile, probs=0.95),
        "size" = colMeans(sizes),
        "size_sd" =  apply(sizes, 2, sd),
        "method" = c(1,2,3),
        "n" = rep(n, 3),
        "e" = rep(e, 3),
        "pi" = rep(pi, 3)
      )
      df = rbind(df, sub_df)
    }
  }
}
```

The results from running the above algorithm are saved in the results_on_recovery_rates.csv file. 

```{r}
recovery_results<- read_csv("results_on_recovery_rates.csv")
```

We generate the plots shown in the paper using these results. 

```{r}
recovery_results %>%
  mutate(method = factor(method, levels = c(1, 2, 3),
                         labels = c("Baseline (CUPED)", "Baseline (Lin)", "Our Approach"))) %>%
  ggplot(aes(x = e, y = recovery, color = method)) +
    geom_line() +
    geom_point() +
    facet_grid(rows = vars(pi), cols = vars(n), 
               labeller = label_both) +
    labs(
      title = "Recovery Rate",
      x = expression("e (Magnitude of Treatment Effect: " * 5 * n^{e} * ")"),
      y = "Recovery Rate",
      color = "Approach"
    ) +
    scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) + 
    scale_color_manual(values = c(
      "Baseline (CUPED)" = "orangered2",
      "Baseline (Lin)" = "orange1",
      "Our Approach" = "dodgerblue3"
    )) +
    theme_bw() +
    theme(legend.position = "bottom")
```