---
title: "Simulations on Powers"
output: github_document
---

```{r setup, include=FALSE}
setwd("/Users/yujin/Documents/identifying-sparse-treatment-effects")
library(glmnet)
library(stabs)
library(readr)
library(dplyr)
library(tidyverse)
```

This function generates data $\{(X_i, T_i, Y_i)\}_{i=1}^n$ given the sample size $n$, the pre-treatment covariate dimension $m$, the outcome dimension $d$, and the treatment assignment probability $p$. 

```{r}
gen_data <- function(n, m, d, p, alphas, sigma = 1, random_seed=0){
  set.seed(random_seed)
  s = length(alphas)
  # randomly generate model parameters beta and delta
  beta = t(replicate(d, runif(m, min=-1, max=1)))
  delta = t(replicate(s, runif(m)))
  treatment = rbinom(n, 1, p)
  # randomly generate X and Y
  X = matrix(rnorm(n*m, sd=1), nrow=n, ncol=m)
  Y = matrix(0, nrow=n, ncol=d)
  for (i in 1:d){
    if (i <= s) {
      Y[,i] =  X %*% beta[i, ] + treatment * (alphas[i] + X %*% delta[i, ]) + 
        rnorm(n, sd=sigma)
    } else {
      Y[,i] =  X %*% beta[i, ] + rnorm(n, sd=sigma)
    }
  }
  return(list("X" = X, "Y" = Y, "treatment" = treatment))
}
```

This function performs subset selection using a baseline ranking-based approach. For more details, see the accompanying file simulations_on_recovery_rates.Rmd. The additional parameter 's_min' specifies the minimum number of dimensions to be selected during subset selection.

```{r}
base_method <- function(X, Y, treatment, method='LIN', s_min=1, s_max=10, cutoff=0.5, B=100){
  
  base_method_helper <- function(X, Y, treatment, alpha=NULL, method='CUPED'){
    
    d = ncol(Y)
    
    if (method == 'CUPED'){
      ## Using CUPED estimator
      lm.fit = lm(Y~X) 
      residuals = lm.fit$residuals
      Y_control = residuals[treatment == 0, ]
      Y_treated = residuals[treatment == 1, ]
      cuped = sapply(1:d, function(i){t.test(Y_treated[,i], Y_control[,i], var.equal=FALSE)$statistic})
      rank_cuped = order(abs(cuped), decreasing = TRUE)
      return (rank_cuped)
    }
    
    if (method == 'LIN') {
      ## Using Lin's estimator
      lm.fit.0 = lm(Y~X, subset = (treatment == 0))
      lm.fit.1 = lm(Y~X, subset = (treatment == 1))
      fitted.0 = predict(lm.fit.0, newdata = as.data.frame(X))
      fitted.1 = predict(lm.fit.1, newdata = as.data.frame(X))
      residuals = Y - mean(treatment) * fitted.0 - mean(1-treatment) * fitted.1
      Y_control = residuals[treatment == 0, ]
      Y_treated = residuals[treatment == 1, ]
      lin = sapply(1:d, function(i){t.test(Y_treated[,i], Y_control[,i], var.equal=FALSE)$statistic})
      rank_lin = order(abs(lin), decreasing = TRUE)
      return (rank_lin)
    }
  }
  
  S = matrix(0, nrow=B, ncol=s_max)
  for (i_b in 1:B){
    indices = sample(1:length(treatment), length(treatment), replace=TRUE)
    X_b = X[indices, ]
    Y_b = Y[indices, ]
    t_b = treatment[indices]
    S[i_b,] = base_method_helper(X_b, Y_b, t_b, method=method)[1:s_max]
  }
  PI = sapply(1:ncol(Y), function(i){sum(S==i)/B})
  selected_set = which(PI>cutoff)
  if (length(selected_set)<s_min) {
    selected_set = order(PI, decreasing=TRUE)[1:s_min]
  }
  return(selected_set)
}
```

This function performs subset selection using our proposed approach. For more details, see the accompanying file simulations_on_recovery_rates.Rmd. The additional parameter 's_min' specifies the minimum number of dimensions to be selected during subset selection.

```{r}
our_method <- function(X, Y, treatment, s_min=1, s_max=10, cutoff=0.5){
  penalty = rep(1, ncol(X) + ncol(Y))
  penalty[(ncol(Y)+1):(ncol(X)+ncol(Y))] = 0 # no penalty on X
  weights = (length(treatment)/sum(treatment))^2 * treatment
  weights = weights + (length(treatment)/sum(1-treatment))^2 * (1-treatment)
  stabsel_res = stabsel(x = scale(cbind(Y, X), center = TRUE, scale = FALSE) * sqrt(weights), 
                        y = (treatment-mean(treatment)) * sqrt(weights),
                        fitfun = glmnet.lasso,
                        args.fitfun = list(family = "gaussian",
                                           alpha = 1, 
                                           nlambda = 200,
                                           penalty.factor = penalty, 
                                           intercept = FALSE, 
                                           standardize = FALSE),
                         cutoff=max(cutoff, 0.6), q=ncol(X)+s_max)
  selected_set = which(stabsel_res$max[1:ncol(Y)]>=cutoff)
  if (length(selected_set)<s_min){
    selected_set = order(stabsel_res$max[1:ncol(Y)], decreasing=TRUE)[1:s_min]
  }
  return(selected_set)
}
```

The following function performs multiple testing and returns significant dimensions. 

```{r}
multiple_testing <- function(X, Y, treatment){
  
  ## Using Lin's estimator
  lm.fit.0 = lm(Y~X, subset = (treatment == 0))
  lm.fit.1 = lm(Y~X, subset = (treatment == 1))
  fitted.0 = predict(lm.fit.0, newdata = as.data.frame(X))
  fitted.1 = predict(lm.fit.1, newdata = as.data.frame(X))
  residuals = Y - sum(treatment)/length(treatment) * fitted.0 - sum(1-treatment)/length(treatment) * fitted.1
  Y_control = residuals[treatment == 0, ]
  Y_treated = residuals[treatment == 1, ]
  
  d = ncol(Y)
  pvalues = sapply(1:ncol(Y), function(i){
    min(t.test(Y_treated[,i], Y_control[,i], var.equal=FALSE)$p.value*d, 1)})
  
  return (which(pvalues<0.05))
}
```

The following function performs simultaneous inference and returns significant dimensions. 

```{r}
psi <- function(X, Y, treatment, B=1000){
  
  ## Get Influence Function
  lm.fit.0 = lm(Y~X, subset = (treatment == 0))
  lm.fit.1 = lm(Y~X, subset = (treatment == 1))
  fitted.0 = predict(lm.fit.0, newdata = as.data.frame(X))
  fitted.1 = predict(lm.fit.1, newdata = as.data.frame(X))
  Y_tilde = Y - mean(treatment) * fitted.0 - mean(1-treatment) * fitted.1
  psi = treatment * Y_tilde / mean(treatment) - (1-treatment) * Y_tilde / mean(1-treatment)
  psi_scaled = scale(psi, center=TRUE, scale=TRUE)
  
  set.seed(0)
  n = nrow(X)
  E = matrix(rnorm(n*B), nrow=n, ncol=B)
  S = (t(psi_scaled) %*% E)/n
  S = apply(S, 2, function(v) max(abs(v)))
  q = quantile(S, 0.95)
  
  tau = colMeans(scale(psi, center=FALSE, scale=TRUE))
  return(which(abs(tau) > q))
}
```

The following functions performs multi-sample splitting procedure with a choice of subset selection method. Set 'func = base_method' to use the baseline ranking-based approach, or 'func = our_method' to use the proposed Lasso-based method. The arguments 's_min', 's_max', and 'cutoff' are passed to the chosen subset selection method. The argument 'p2' specifies the proportion of the sample allocated to the second split for effect estimation, and B is the number of multi-splits used in the procedure. The function outputs significant dimensions.

```{r}
multi_sample_splitting <- function(X, Y, treatment, func,
                                   s_min=1, s_max=10, cutoff=0.5, p2=0.6, B=100){
  n = nrow(X); d = ncol(Y)
  n1=n*(1-p2)
  
  evaluate_on_the_second_split <- function(X, Y, treatment, selected_set){
    
    lm.fit.0 = lm(Y~X, subset = (treatment == 0))
    lm.fit.1 = lm(Y~X, subset = (treatment == 1))
    fitted.0 = predict(lm.fit.0, newdata = as.data.frame(X))
    fitted.1 = predict(lm.fit.1, newdata = as.data.frame(X))
    residuals = Y - mean(treatment) * fitted.0 - mean(1-treatment) * fitted.1
    Y_control = residuals[treatment == 0, ]
    Y_treated = residuals[treatment == 1, ]
  
    pvalues = rep(1, ncol(Y))
    for (s in selected_set){
      pvalues[s] = t.test(Y_treated[,s], Y_control[,s], var.equal=FALSE)$p.value * length(selected_set)
    }
    return (pvalues)
  }
  
  pvalues = matrix(1, nrow = B, ncol = d)
  for (i_b in 1:B){
    set.seed(i_b)
    indices = sample(1:n, n, replace=FALSE)
    first_split = indices[1:n1]; second_split = indices[(n1+1):n]
    selected_set = func(X[first_split, ], Y[first_split, ], 
                        treatment[first_split], 
                        s_min=s_min, s_max=s_max, cutoff=cutoff)
    pvalues[i_b,] = evaluate_on_the_second_split(X[second_split, ], Y[second_split, ], 
                                                 treatment[second_split], 
                                                 selected_set = selected_set)
  }
  selected_set_f = which(apply(pvalues, 2, function(x) {min(quantile(x, 0.1) * 10, 1)}) < 0.05)
  return (selected_set_f)
}
```

The following code runs simulations to compare powers of the following methods: 1) multiple testing, 2) simultaneous inference, 3) multi-sample splitting with baseline ranking-based subset selection method, and 4) multi-sample splitting with our subset selection method using the Lasso. 

```{r, eval=FALSE}
d = 500; m = 50; B = 50

df = data.frame()
n_vec = c(500)
pi_vec = c(0.3, 0.5, 0.7)
alphas = c(1, 1, 1, -1, -1)
exp_vec = c(-1/2.5, -1/2.75, -1/3, -1/3.25, -1/3.5, -1/3.75, -1/4)
exp_vec = c(-1/3.5)
for (pi in pi_vec){
  for (n in n_vec) {
    for (e in exp_vec){
      alpha = 5*(n^(e))
      res = matrix(0, nrow=B, ncol=4)
      sizes = matrix(0, nrow=B, ncol=4)
      for (i_b in 1:B){
        data = gen_data(n, m, d, pi, alphas=alpha*alphas, sigma=1, random_seed=i_b)
        Y = data$Y; X = data$X; treatment = data$treatment
        s1 = multiple_testing(X, Y, treatment)
        s2 = psi(X, Y, treatment)
        s3 = multi_sample_splitting(X, Y, treatment, base_method, cutoff=0.4)
        s4 = multi_sample_splitting(X, Y, treatment, our_method, cutoff=0.4)
        res[i_b, ] = c(sum(s1 %in% c(1,2,3,4,5))/5, sum(s2 %in% c(1,2,3,4,5))/5, 
                       sum(s3 %in% c(1,2,3,4,5))/5, sum(s4 %in% c(1,2,3,4,5))/5)
        sizes[i_b, ] = c(length(s1), length(s2), length(s3), length(s4))
      }
      sub_df = data.frame(
        "power" = colMeans(res),
        "method" = c(1,2,3,4),
        "n" = rep(n, 4),
        "e" = rep(e, 4),
        "pi" = rep(pi, 4)
      )
      df = rbind(df, sub_df)
    }
  }
}
```

The results from running the above algorithm are saved in the results_on_powers.csv file. 

```{r}
power_results<- read_csv("results_on_powers.csv")
```

We generate the plots shown in the paper using these results. 

```{r}
ggplot(power_results, aes(x = e, y = power, color = method, linewidth=method)) +
  geom_line() +
  geom_point() +
  facet_grid(rows = vars(n), cols = vars(pi),
             labeller = label_both) +
  labs(
    x = expression("e (Magnitude of Treatment Effect: " * 5 * n^{e} * ")"),
    y = "True Positive Rate",
    color = "Method",
  ) +
  scale_linewidth_manual(values = c(
    "Multiple Testing" = 2,
    "Simultaneous Inference" = 1,
    "Multi-Sample Splitting\nwith Baseline Approach\nfor Subset Selection" = 1,
    "Multi-Sample Splitting\nwith Our Approach\nfor Subset Selection" = 1
  )) +
  scale_color_manual(values = c(
    "Multiple Testing" = "orangered2",
    "Simultaneous Inference" = "orange1",
    "Multi-Sample Splitting\nwith Baseline Approach\nfor Subset Selection" = "forestgreen",
    "Multi-Sample Splitting\nwith Our Approach\nfor Subset Selection" = "dodgerblue3"
  )) +
  guides(linewidth = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05))) +
  theme_bw() +
  theme(
    legend.position = "right",
    legend.spacing.y = unit(8, "pt"),  # Increase vertical space between legend items
    legend.text = element_text(size=10, margin = margin(b = 6)),  # Optional: extra bottom margin
    axis.title.x = element_text(margin = margin(t = 14)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
